### 2024-09-25 paper/paper.qmd

---
title: "Outpacing the Lanes: Toronto's Bike Share Ridership Surges Ahead of Infrastructure Growth"
author: 
  - Steven Li
thanks: "Code and data are available at: https://github.com/stevenli-uoft/Toronto_BikeShare_Development"
date: today
date-format: long
abstract: "Toronto's Bike Share program experienced a nearly fivefold increase in annual rides from 2017 to 2023, outpacing the growth of the city's cycling infrastructure. Analysis of open data reveals that while bikeways increased from 539 in 2001 to over 1,400 by 2023, this expansion rate lags behind the surge in ridership. The mismatch between infrastructure development and Bike Share usage highlights an urgent need for accelerated and strategic expansion of Toronto's cycling network. These findings underscore the importance of data-driven urban planning to ensure safe, accessible cycling options for a growing user base."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(knitr)
library(lubridate)
library(ggplot2)

set.seed(778)
```

\newpage

# Introduction
Cities worldwide, grappling with congestion and pollution, are turning to cycling as a sustainable transportation solution. Toronto, Canada's largest city, has been at the forefront of this trend, implementing a bike-sharing program and expanding its cycling infrastructure to encourage more residents to adopt this eco-friendly mode of transport.

Toronto's Bike Share program, launched in 2011, has become an integral part of
the city's transportation network. Managed by Bike Share Toronto under the 
Toronto Parking Authority, the program offers thousands of bicycles at hundreds
of stations throughout the city, providing an accessible and affordable option 
for short trips [@bikeshareToronto]. The system has grown significantly since 
its inception, recently expanding to include e-bikes and integrating with the 
city's public transit system [@ttcIntegration].

Concurrent with the growth of the Bike Share program, Toronto has been 
investing in its cycling infrastructure. The city's Cycling Network Plan 
outlines key improvements and expansions to bike lanes, trails, and shared 
routes [@cyclingNetworkPlan]. This includes the introduction of dedicated bike
lanes in high-traffic areas, the development of multi-use trails, and the 
continuous expansion of the cycling network to improve connectivity across 
the city.

Several key initiatives have marked Toronto's commitment to cycling 
infrastructure. The Bloor-Danforth Pilot project in 2016 demonstrated the 
viability of expanded cycling lanes on major roads [@bloorDanforthPilot]. 
The ActiveTO program, initiated in response to the COVID-19 pandemic, 
temporarily closed major streets to car traffic, leading to some bike lanes 
becoming permanent fixtures [@activeTO]. More recently, the city has committed 
to adding 100 kilometers of new cycling infrastructure between 2022 and 2024, 
with plans to complete 500 kilometers by 2030 [@cyclingPlan2030].

As Toronto continues to invest in both its Bike Share program and cycling 
infrastructure, an important question arises: Is the city's bikeway 
infrastructure keeping pace with the growth in bike share ridership? This
paper aims to explore this question by analyzing data on Toronto's Bike Share 
ridership and the development of its cycling infrastructure from 2017 to 2023.

The remainder of this paper is structured as follows. @sec-data describes the 
data sources and methodology used in our analysis. @sec-results presents the 
results of our data exploration, including trends in Bike Share ridership and
the evolution of Toronto's cycling infrastructure. @sec-discussion discusses 
the implications of these findings, addressing the relationship between
infrastructure development and ridership growth. Finally, @sec-limit concludes
with limitations and suggestions for future research and policy considerations.

# Data {#sec-data}
```{r}
#| echo: false
#| warning: false
#| message: false

bikeShare_data <- read_csv(here::here(
  "data/analysis_data/monthly_aggregated_bikeshare_data.csv"))

bikeWay_data <- read_csv(here::here("data/analysis_data/bikeway_data.csv"))

bikeway_colors <- c("Shared Roadways" = "#e15759", 
                    "On-Road Lanes" = "#76b7b2",
                    "Protected Lanes" = "#2ca02c")
```

This study utilizes two primary datasets: Toronto's Bike Share ridership data 
[@bikeShareData] and the city's Cycling Network (Bikeways) data 
[@bikeWayData]. Both datasets were obtained from the City of Toronto's Open Data Portal [@torontoOpenData], providing a comprehensive view of cycling infrastructure from 2001 to 2024, and Bike Share ridership in the city from 2017 to 2023.

## Bike Share Ridership Data
The Bike Share ridership data was collected by Bike Share Toronto, a division of the Toronto Parking Authority. The raw data consisted of anonymized trip data, with each month or quarter provided as a separate CSV file containing all unique trips. A "ride" in this context is defined as a journey that starts at one station and ends at another.

We cleaned the data by downloading all CSV files, extracting required columns (trip ID and start date), and performing necessary cleaning operations. This process involved standardizing column names across files, converting data types, and aggregating the data to calculate monthly total ridership figures. The cleaned data provides a comprehensive view of Bike Share usage from January 2017 to December 2023.

One challenge in processing this data was the inconsistency in column names and data formats across different years. This was addressed through a custom R function that standardized column names and data types across all files. Additionally, any incomplete or invalid entries were removed during the cleaning process to ensure data integrity.

A sample of this processed dataset can be seen in (@tbl-sample-bikeshare), showcasing the year and month of observation along with the total number of rides for that month. This granular monthly data allows for analysis of seasonal trends and year-over-year growth in ridership.

```{r}
#| label: tbl-sample-bikeshare
#| tbl-cap: "Sample of Toronto's Bike Share Monthly Ridership Data (2017-2023)"
#| echo: false
#| warning: false
#| message: false

bikeShare_sample <- bikeShare_data %>% sample_n(5) %>% 
  as_tibble() %>% arrange(start_date)

colnames(bikeShare_sample) <- c("Year & Month", "Total Monthly Rides")

kable(bikeShare_sample, format.args = list(big.mark = ","))
```

To provide an overview of the Bike Share program's growth, we aggregated the monthly data into annual statistics, as shown in (@tbl-ridership_summary_stats).

```{r}
#| label: tbl-ridership_summary_stats
#| tbl-cap: "Annual Bike Share Ridership Summary Statistics"
#| echo: false
#| warning: false
#| message: false

bikeShare_summary <- bikeShare_data %>%
  mutate(year = year(start_date)) %>%  
  group_by(year) %>%   
  summarise(
    total_rides = sum(total_rides),
    avg_monthly_rides = mean(total_rides)
  ) %>%
  mutate(
    year_on_year_growth = (total_rides - lag(total_rides)) / lag(total_rides) * 100
  ) %>%
  mutate(
    year_on_year_growth = round(year_on_year_growth, 2),  # Round to 2 decimal places
    total_rides = format(total_rides, big.mark = ","),  # Add comma separators
    avg_monthly_rides = format(round(avg_monthly_rides), big.mark = ",")  # Round and add comma separators
  )

colnames(bikeShare_summary) <- c("Year", "Total Rides", "Avg Monthly Rides", 
                                 "Year-on-year Growth (%)")

kable(bikeShare_summary)
```

This summary reveals a consistent increase in total annual rides from 1,190,320 in 2017 to 5,713,141 in 2023, representing nearly a fivefold increase over seven years. The average monthly rides have grown correspondingly, indicating a substantial expansion of the program's usage.

## Cycling Network (Bikeways) Data
The Cycling Network data was sourced from the City of Toronto's Transportation Services division. This dataset provides detailed information about the city's bikeways, including their construction year, any upgrades, and the type of bikeway. (@tbl-sample_bikeway) offers a sample of this dataset.

```{r}
#| label: tbl-sample_bikeway
#| tbl-cap: "Sample of Toronto's Cycling Network (Bikeways) Data (2001-2023)"
#| echo: false
#| warning: false
#| message: false

bikeWay_sample <- bikeWay_data %>% sample_n(5) %>% 
  as_tibble() %>% arrange(INSTALLED)

colnames(bikeWay_sample) <- c("Route ID", "Constructed Year", "Upgraded Year"
                              , "Bike Way Type")

kable(bikeWay_sample)
```

The raw data contained more detailed specifications for bikeway types, which we grouped into three main categories for ease of understanding and analysis:

1. **Protected Lanes**: Physically separated and protected from motor vehicle traffic. This includes Bi-Direction Cycle Tracks, Cycle Tracks, and Multi-Use Trails.
2. **On-Road Lanes**: Designated lanes on the road, but not physically separated. This category encompasses various types of bike lanes.
3. **Shared Roadways**: Roads where cyclists and motor vehicles share the same space. This includes Sharrows, Signed Routes, and Park Roads.

It's important to note that in the original dataset, bikeways constructed before 2001 were all labeled with a construction year of 2001. This data limitation could potentially lead to skewing or misunderstanding in the analysis, particularly for earlier years. Additionally, some entries with invalid years (e.g., year labeled as "1") or missing data were filtered out during the cleaning process to ensure data quality.

Another key limitation to note about the bikeway data, is that our analysis relies on the count of bikeways rather than their total length. This approach was necessitated by data availability constraints, as information on the length of each bikeway was not provided in the dataset. Using counts instead of lengths may not provide a fully accurate representation of infrastructure growth, as it doesn't account for variations in bikeway size or capacity. For instance, a short protected lane would be given the same weight in our analysis as a long multi-use trail. This limitation should be kept in mind when interpreting the results, particularly when comparing infrastructure growth to ridership increases.

Both datasets were cleaned and processed using R [@RCoreTeam2023], with additional support from the tidyverse [@tidyverse], dplyr [@dplyr], and lubridate [@lubridate] packages. The cleaning process involved removing any incomplete entries, standardizing date formats for consistency across the datasets, and aggregating data where necessary.

These datasets, chosen for their up-to-date and comprehensive nature, offer the most accurate available information on Toronto's cycling infrastructure and Bike Share usage. When analyzed together, they provide insights into the relationship between infrastructure development and the growing demand for cycling in Toronto, allowing us to assess whether the city's bikeway expansion is keeping pace with increasing ridership. While other datasets were considered, such as bike lanes measured in kilometers or total bike counts, they were not used due to being outdated or no longer maintained.

# Results {#sec-results}
Our analysis of Toronto's Bike Share ridership and cycling infrastructure data from 2017 to 2023 reveals several key trends and patterns, shedding light on the relationship between infrastructure development and cycling adoption in the city.

## Bike Share Ridership Trends
Toronto's Bike Share monthly ridership from 2017 to 2023 is illustrated in (@fig-ridership_monthly_graph), showing a clear upward trajectory with pronounced seasonal fluctuations. Peak usage consistently occurs during the summer months, while winter sees a significant reduction in rides. Despite these seasonal variations, the overall trend line demonstrates steady growth in ridership.

The most notable features of this trend include:

1. A consistent year-over-year increase in peak monthly ridership, indicating growing popularity of the program.
2. Gradual expansion of the riding season, with shoulder months (spring and fall) showing increased activity over time, suggesting a broadening user base.
3. A brief dip in ridership during early 2020, likely attributable to initial COVID-19 lockdowns, followed by a robust recovery. This recovery may be partially attributed to the ActiveTO program, which closed major streets to car traffic, creating more space for cyclists [@activeTO].

```{r}
#| label: fig-ridership_monthly_graph
#| fig-cap: "Toronto Bike Share Monthly Ridership Trend (2017-2023)"
#| echo: false
#| warning: false
#| message: false

ggplot(bikeShare_data, aes(x = start_date, y = total_rides)) +
     geom_line(aes(color = "Monthly Ridership"), size = 1) +
     geom_smooth(aes(color = "Trendline"), method = "loess", se = FALSE) +
     labs(
       x = "Date",
       y = "Total Rides",
       color = NULL
     ) +
     theme_minimal() +
     scale_color_manual(values = c("Monthly Ridership" = "#1f77b4", 
                                   "Trendline" = "#ff7f0e")) +
     scale_x_date(
       date_breaks = "1 year", 
       date_labels = "%Y",
       limits = c(as.Date("2017-01-01"), as.Date("2023-12-31")),
       expand = c(0, 0)
     ) + 
     scale_y_continuous(
       labels = function(x) paste0(x/1000, "k"),
       breaks = seq(0, max(bikeShare_data$total_rides), by = 150000),
       minor_breaks = seq(0, max(bikeShare_data$total_rides), by = 75000)
     ) +
     theme(
       legend.position = "top",
       panel.grid.major = element_line(color = "gray85", linewidth = 0.35)
     )
```

## Cycling Infrastructure Development
A breakdown of bikeway types constructed each year from 2002 to 2023 is 
presented by (@fig-bikeway_type_constructed_graph). This graph reveals several 
interesting patterns in the development of Toronto's cycling infrastructure:

1. A significant spike in bikeway installations occurred between 2005 and 2007, primarily driven by the addition of shared roadways. This surge aligns with the city's efforts to implement its 2001 Bike Plan, which aimed to create 1,000 km of bikeways by 2011. While the city fell short of this ambitious target, the period saw a substantial increase in shared roadways, adding over 145 km by 2007 [@ratchford]. As shown in (@fig-cumulative-proportion-bikeways), this rapid expansion of shared roadways significantly altered the proportional makeup of Toronto's bikeway network, with shared roadways dominating the infrastructure mix for several years following 2007.
2. After 2007, there's a noticeable shift towards more protected lanes and on-road lanes, reflecting growing concerns about cyclist safety and calls for more separated cycling infrastructure.
3. The installation of new bikeways has been relatively consistent from 2010 
onwards, with a mix of all three types (shared roadways, on-road lanes, and 
protected lanes) being added each year.

```{r}
#| label: fig-bikeway_type_constructed_graph
#| fig-cap: "Breakdown of Bikeway Types by Installation Year (2002-2023)"
#| fig-foot: "Bikeways installed prior to 2001 are shown with an install year of 2001"
#| echo: false
#| warning: false
#| message: false

bikeWay_type_summary <- bikeWay_data %>%
  filter(INSTALLED >= 2002 & INSTALLED <= 2023) %>%
  group_by(INSTALLED, INFRA_HIGHORDER) %>%
  summarise(count = n())

ggplot(bikeWay_type_summary, aes(x = as.factor(INSTALLED), y = count, fill = INFRA_HIGHORDER)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = bikeway_colors) +
  labs(
    x = "Year Installed",
    y = "Count of Bikeways",
    fill = "Bikeway Types:"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray85", linewidth=0.35),
    legend.position = "top"
  )
```

Bikeways were also upgraded throughout the years to improve the quality and
safety for bikers.(@fig-bikeway_type_upgraded_graph) shows the breakdown of 
bikeway upgrades by year from 2002 to 2023. Notable observations include:

1. A significant spike in upgrades around 2015, primarily focused on converting existing bikeways to protected lanes. This likely reflects the city's response to increasing demands for safer cycling infrastructure and aligns with several initiatives that began development around that time.
2. After 2015, upgrades have continued at a lower but steady rate, with a focus on improving existing infrastructure rather than just adding new bikeways.

```{r}
#| label: fig-bikeway_type_upgraded_graph
#| fig-cap: "Breakdown of Bikeway Types by Upgraded Year (2002-2023)"
#| echo: false
#| warning: false
#| message: false

bikeWay_type_summary <- bikeWay_data %>%
  filter(UPGRADED >= 2002 & UPGRADED <= 2023) %>%
  group_by(UPGRADED, INFRA_HIGHORDER) %>%
  summarise(count = n()) %>%
  drop_na(UPGRADED)

ggplot(bikeWay_type_summary, aes(x = as.factor(UPGRADED), y = count, fill = INFRA_HIGHORDER)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = bikeway_colors) +
  labs(
    x = "Year Upgraded",
    y = "Count of Bikeways",
    fill = "Bikeway Types:"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray85", linewidth=0.35),
    legend.position = "top"
  )
```

## Cumulative Growth of Cycling Infrastructure
To provide a complete understanding of Toronto's entire Cycling network, 
(@fig-cumulative-growth) presents the cumulative growth of different bikeway 
types from 2001 to 2023, demonstrating:

1. A steady increase in the total number of bikeways over the two decades, from 539 in 2001 to 1,446 by 2023.
2. The rapid growth of shared roadways in the mid-2000s, which has since 
stabilized as a proportion of total infrastructure.
3. A consistent growth in protected lanes, particularly accelerating in recent years, aligning with the city's commitment to add 100 kilometers of new cycling infrastructure between 2022 and 2024, and 500 kilometers by 2030 [@cyclingPlan2030].

```{r}
#| label: fig-cumulative-growth
#| fig-cap: "Cumulative Bikeway types by Installed Year (2001-2023)"
#| echo: false
#| warning: false
#| message: false

cumulative_bikeways <- bikeWay_data %>%
  filter(INSTALLED >= 2001 & INSTALLED <= 2023) %>%
  group_by(INSTALLED, INFRA_HIGHORDER) %>%
  summarise(count = n(), .groups = "drop") %>%
  complete(INSTALLED = 2001:2023, INFRA_HIGHORDER, fill = list(count = 0)) %>%
  arrange(INFRA_HIGHORDER, INSTALLED) %>%
  group_by(INFRA_HIGHORDER) %>%
  mutate(cumulative_count = cumsum(count)) %>%
  ungroup()

# Create the bar graph
ggplot(cumulative_bikeways, aes(x = INSTALLED, y = cumulative_count, fill = INFRA_HIGHORDER)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = bikeway_colors) +
  labs(
    x = "Year Installed",
    y = "Cumulative Number of Bikeways",
    fill = "Bikeway Types:"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray85", linewidth = 0.35),
    legend.position = "top"
  ) +
  scale_x_continuous(breaks = seq(2001, 2023, by = 1))
```

\newpage
## Comparison of Ridership and Infrastructure Growth
To directly compare the ridership and infrastructure growth, (@fig-combined_ridership_bikeways_graph) provides a comparative view of annual 
Bike Share ridership and the cumulative number of bikeways from 2017 to 2023.
This graph reveals:

1. A strong positive correlation between the growth in Bike Share ridership 
and the expansion of cycling infrastructure.
2. Bike Share ridership has grown at a faster rate than the expansion of bikeways. However, it's important to note that this comparison is limited by the use of bikeway count rather than total length, which might provide a more accurate representation of infrastructure growth.
3. Despite the apparent slower growth of bikeways compared to ridership, the continuous increase in infrastructure likely supports and encourages the growing popularity of cycling in Toronto.

```{r}
#| label: fig-combined_ridership_bikeways_graph
#| fig-cap: "Comparison of Annual Bike Share Ridership and Cumulative Bikeways (2017-2023)"
#| echo: false
#| warning: false
#| message: false

# Summarize cumulative bikeway data for the years 2017-2023
bikeWay_cumulative_summary <- bikeWay_data %>%
  filter(INSTALLED <= 2023) %>%
  group_by(INSTALLED) %>%
  summarise(total_bikeways_installed = n()) %>%
  mutate(cumulative_bikeways = cumsum(total_bikeways_installed)) %>%
  filter(INSTALLED >= 2017)

# Summarize bike share data for the years 2017-2023
bikeShare_yearly_summary <- bikeShare_data %>%
  mutate(year = year(start_date)) %>%
  filter(year >= 2017 & year <= 2023) %>%
  group_by(year) %>%
  summarise(total_rides = sum(total_rides))

# Merge the two datasets by year
combined_data <- left_join(bikeShare_yearly_summary, 
                           bikeWay_cumulative_summary, 
                           by = c("year" = "INSTALLED"))

ggplot(combined_data, aes(x = year)) +
  geom_bar(aes(y = cumulative_bikeways * 2000, fill = "Bikeways"), 
           stat = "identity", alpha = 0.75) +
  geom_line(aes(y = total_rides, color = "Ridership"), size = 1) +
  scale_y_continuous(
    name = "Total Annual Rides", 
    sec.axis = sec_axis(~ . / 2000, name = "Cumulative Number of Bikeways"),
    labels = function(x) paste0(x/1000000, "M"),
    breaks = seq(0, max(combined_data$total_rides), by = 1000000),
    minor_breaks = seq(0, max(combined_data$total_rides), by = 500000)
  ) +
  scale_x_continuous(
    breaks = seq(2017, 2023, by = 1),
    limits = c(2016.5, 2023.5),
    expand = c(0, 0)
  ) +
  scale_color_manual(values = c("Ridership" = "#1f77b4")) +
  scale_fill_manual(values = c("Bikeways" = "gray50")) +
  labs(
    x = "Year",
    color = NULL,
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.major = element_line(color = "gray85", linewidth=0.35)
  )
```

These results suggest that while Toronto has made significant strides in expanding its cycling infrastructure, the growth in Bike Share ridership may be outpacing the development of new bikeways. This trend underscores the importance of continued investment in cycling infrastructure to meet the growing demand for safe and accessible cycling options in the city.

# Discussion {#sec-discussion}
The analysis reveals a complex interplay between infrastructure investment and cycling adoption in Toronto from 2017 to 2023. This discussion will explore the implications of our findings, addressing the central question of whether Toronto's bikeway infrastructure is keeping pace with the growth in Bike Share ridership.

## Rapid Growth in Bike Share Utilization
The most striking finding from our analysis is the dramatic increase in Bike Share ridership over the study period. The nearly fivefold increase in annual rides from 2017 to 2023 demonstrates a significant shift in urban mobility patterns and a growing acceptance of cycling as a viable transportation option in Toronto. This trend aligns with global movements towards more sustainable and active transportation modes in urban areas.

The consistent year-over-year growth, even in the face of the COVID-19 pandemic, suggests a resilient and potentially permanent change in transportation habits among Toronto residents. This rapid adoption of Bike Share presents both opportunities and challenges for urban planners and policymakers.

## Infrastructure Development: Progress and Challenges
Our analysis of cycling infrastructure development reveals a more nuanced picture. While there has been steady growth in the total number of bikeways, from approximately 600 in 2001 to over 1,400 by 2023, the rate of infrastructure expansion has not matched the pace of Bike Share ridership growth. This disparity raises questions about the capacity of the current infrastructure to safely and efficiently accommodate the increasing number of cyclists.

The shift in focus from shared roadways to more protected lanes, particularly evident after 2007, reflects an evolving understanding of cyclist safety needs. This aligns with the city's commitment to Vision Zero and Complete Streets policies, which prioritize safety for all road users [@cyclingNetworkPlan]. However, the data suggests that there may be a need for accelerated implementation of these safer infrastructure options to meet the growing demand.

## Policy Implications and Future Directions
The rapid growth in Bike Share ridership, outpacing infrastructure development, underscores the need for more aggressive expansion of Toronto's cycling network. The city's recent commitment to adding 100 kilometers of new cycling infrastructure between 2022 and 2024, with plans to complete 500 kilometers by 2030 [@cyclingPlan2030], is a step in the right direction. However, our findings suggest that even this ambitious plan may need to be accelerated or expanded to meet the growing demand.
Several policy implications emerge from our analysis:

1. Prioritizing Protected Infrastructure: Given the clear preference for safer cycling options, future infrastructure development should prioritize protected lanes over shared roadways where feasible.
2. Strategic Expansion: The city should consider strategically expanding the Bike Share program in tandem with new infrastructure development to ensure a balanced growth of both supply and demand.
4. Integration with Public Transit: The success of the Bike Share program suggests opportunities for further integration with other forms of public transportation, potentially addressing last-mile connectivity issues.

## Limitations and Future Research {#sec-limit}
While our study provides valuable insights into the relationship between cycling infrastructure and Bike Share usage in Toronto, it has several limitations that suggest avenues for future research:

1. **Geographic Distribution**: This study did not examine the spatial distribution of Bike Share usage and infrastructure. Future research could employ geographic information to map these distributions, revealing patterns of usage and infrastructure gaps.
2. **Quality of Infrastructure**: Our analysis focused on infrastructure quantity rather than quality or connectivity. Future studies could incorporate these measures for a more comprehensive understanding of Toronto's cycling landscape.
3. **Infrastructure Measurement**: Due to data limitations, our analysis relied on the count of bikeways rather than their length. Future studies with access to length data could offer a more precise comparison between infrastructure development and ridership growth, potentially creating a weighted index that accounts for factors such as bikeway type, length, and connectivity.
4. **Comparative Analysis**: A study comparing Toronto with cities of similar climate and urban structure could help identify best practices in balancing infrastructure development with growing cycling demand.

These research directions could significantly enhance our understanding of the complex relationship between cycling infrastructure development and ridership growth in urban environments.

In conclusion, while Toronto has made significant strides in expanding its cycling infrastructure and Bike Share program, our analysis suggests that infrastructure development is lagging behind the rapid growth in cycling demand. As the city continues to invest in sustainable transportation options, a data-driven, strategic approach to cycling infrastructure development will be crucial in supporting and encouraging the growing cycling culture in Toronto.

\newpage

\appendix

# Appendix {-}

The following graph shows how the composition of Toronto's bikeway network has evolved over time, taking into account the cumulative additions of different bikeway types.

```{r}
#| label: fig-cumulative-proportion-bikeways
#| fig-cap: "Proportional Breakdown of Cumulative Bikeways by Type and Year (2001-2023)"
#| echo: false
#| warning: false
#| message: false

# Calculate cumulative counts for each bikeway type by year
cumulative_bikeways <- bikeWay_data %>%
  filter(INSTALLED >= 2001 & INSTALLED <= 2023) %>%
  group_by(INSTALLED, INFRA_HIGHORDER) %>%
  summarise(count = n(), .groups = "drop") %>%
  complete(INSTALLED = 2001:2023, INFRA_HIGHORDER, fill = list(count = 0)) %>%
  arrange(INFRA_HIGHORDER, INSTALLED) %>%
  group_by(INFRA_HIGHORDER) %>%
  mutate(cumulative_count = cumsum(count)) %>%
  ungroup()

# Calculate the total cumulative count for each year
total_cumulative <- cumulative_bikeways %>%
  group_by(INSTALLED) %>%
  summarise(total = sum(cumulative_count))

# Join the total to the main dataset and calculate proportions
cumulative_proportions <- cumulative_bikeways %>%
  left_join(total_cumulative, by = "INSTALLED") %>%
  mutate(proportion = cumulative_count / total)

# Create the proportional stacked bar plot
ggplot(cumulative_proportions, aes(x = INSTALLED, y = proportion, fill = INFRA_HIGHORDER)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = bikeway_colors) +
  labs(
    x = "Year",
    y = "Proportion of Cumulative Bikeways",
    fill = "Bikeway Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  scale_x_continuous(breaks = seq(2001, 2023, by = 2))
```

This graph provides additional context to the absolute numbers shown in the main text. It illustrates how the relative proportions of different bikeway types have changed as the network has grown, which may reflect shifts in urban planning priorities or infrastructure strategies over the years.

\newpage
# References {-}